The GitHub Open Source Development Process \author{Kevin Peterson\\ Department of Information Technology\\ Clinical Informatics Support Systems\\ Mayo Clinic\\ 200 1st Street\\ Rochester, MN 55905\\ \small \texttt{peterson.kevin@mayo.edu} }

### 2013-04-11 paper/main.tex
First run through proof-reading.

\documentclass{proc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{fixltx2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{url}
\usepackage{amsmath}
%\setlength{\parskip}{\baselineskip}
%\setlength{\parindent}{0pt}

@@

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}. Along with hosting software repositories, GitHub incorporates social aspects of development. GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions. This type of social integration of a user's identity and actions is unique to GitHub\cite{dabbish2012social}.

Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
Git also allows for a variety of work flows\cite{chacon2009pro}. These workflows may be tailored to the individual needs of a project. The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature. Other work flows may be more distributed, or resemble more traditional centralized SCM systems. Although work flow techniques aren't explicitly addressed in this work, it is important to note that work flow change may contribute to \textit{process change}.

@@

In addition, three supporting research questions are proposed. These questions will draw upon data from this work, but also reference interviews with GitHub staff to better represent how GitHub may be changing how OSS development is conducted. These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.

@@

\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository. As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\

@@

The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way commercial software is. ``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}. ``Patches'' in the context of GitHub could be equated to Pull Requests.

This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\

@@

\subsection{Research Questions}
To expand the analysis scope slightly, three qualitative research questions are posed. These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub. Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.\\

\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\

@@

\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The Object Management Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards. Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.

\section{Methods}
\subsection{Collection and Storage}
Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}. A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.

\begin{algorithm}[H]
\captionof{algorithm}{Data Collection Algorithm}\label{algo:data_collection_algo}
\begin{algorithmic}[1]
\While{$i < 1000$} 
\State $i \gets i + 1$
\State $word \gets random(word\_list)$
\State $repos \gets git\_search(word)$
\State $j \gets $%\Call{random}{0... \Call{size}{repos}}$
\State $repo \gets repos_j$
\State \Call{store}{repo}
\EndWhile
\end{algorithmic}
\end{algorithm}

The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis. Results were stored in a MySQL relational database management system. The database schema to store the data is described in Figure \ref{fig:er_diagram}.

@@

All source code and materials used in data collection and analysis are hosted at GitHub (naturally!), and may be accessed by the following URL: \url{https://github.com/kevinpeterson/github-process-research}

\subsection{Analysis}
In this context, a \textit{Repository} is defined as a Set of \textit{Attributes}. An \textit{Attribute} is a metric of interest. See Section \ref{sec:attributes} for further information on individual \textit{Attributes}. These metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:

@@

All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a Set of Sets of \textit{Attributes}. Let the Set of these Sets of \textit{Attribute} Sets be called $S$, or the \textit{Repository Sample}. We can label the \textit{Attribute} Sets by arbitrary integers 0 through $|S|$, or:
\[ \{A_i\}_{i \in \{0,1,2 ... |S|\} }\]

In this way, we can reference a Set of \textit{Attributes} (or, an individual \textit{Repository}), by $A_i$.

@@

\noindent \textbf{\textit{Commits:}}
The number of individual source code commits to a repository. It is important to note that a textit{commit} is a single transaction that changes the code contents of a repository. Actual commit ``size'' (lines of code modified, for example), is not taken into account.

@@

\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for serveral reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests. TODOs, or general support questions. \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.

@@

Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related. The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip. More investigation would be necessary to further examine this finding.\\

\subsection{Research Questions}
\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown. The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or:

\begin{equation}
P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\}
\label{eq:commit_percentage}
\end{equation}

Where $U$ is the Set of \textit{Committers} to a given \textit{Repository} and $C$ is the MultiSet of \textit{Commits} by an individual \textit{Committers} to a given \textit{Repository}, or:
 \[ \{ C_u \}_{u \in U} \]

\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Individual committer commit percentage per repository}
\label{fig:committers_percentage_pie_chart}
\end{figure}

To find the percentages Set $P$ of equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository. Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range). The distribution of this range assignments is plotted on Figure \ref{fig:committers_percentage_pie_chart}.

There does appear to be evidence of a small set of core developers, denoted by the relatively high percentage of developers that have authored $95-100\%$ of the commits of a given repository. Most notably, however, is that the data shows that by a large percentage of developers committed small amounts ($0-5\%$ of the total repository commits). This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways.

@@

When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (figure \ref{fig:committers_histogram}).\\

\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook. In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.

Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for develper trust building:
\begin{quote}
Only in a society where technical supremacy is highly appreciated can developers acquire good reputations among their peers by displaying their skills through free distribution, and often wider acceptance, of their systems.The good reputation attracts attention, trust, and cooperation from others and lays the foundation for advancing the original developers agenda and the establishment and development of OSS communities. 
\end{quote}

\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.

As a case study, the CTS2 OMG\textregistered Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes. Elicitation, Analysis and Negotiation were natural fits for the issue tracker. This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issues tracker itself allowed for dialogs in the form of comments. Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern. OMG has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
Validation and Management are areas of requirements engineering that GitHub has interesting solutions for. First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log. This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue. As the normal OMG process proceeds, requirements can be managed through ``tagging'' the issues. By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own workflow. For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration

As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.

\section{Conclusion}

>>>>

\documentclass{proc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{fixltx2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathrsfs}
%\setlength{\parskip}{\baselineskip}
%\setlength{\parindent}{0pt}

@@

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates social aspects of development. GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions. This type of social integration of a user's identity and actions is unique to GitHub\cite{dabbish2012social}.

Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
Git also allows for a variety of work flows\cite{chacon2009pro}. These work flows may be tailored to the individual needs of a project. The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature. Other work flows may be more distributed, or resemble more traditional centralized SCM systems. Although work flow techniques aren't explicitly addressed in this work, it is important to note that work flow change may contribute to \textit{process change}.

@@

In addition, three supporting research questions are proposed. These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing how OSS development is conducted. These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.

@@

\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.

In GitHub, a \emph{watch} is a user act of subscribing to changes in a repository. It is also referred to as a \emph{star}. A \emph{watch} allows users to receive status updates, such as when code has been committed. This is considered a \emph{passive} act, as it signifies involvement in the community, but no direct action to contribute.

A \emph{fork} is the creation of a personal copy of a repository\cite{dabbish2012social}, from which they can change without impacting the original code source. A \emph{fork} is an \emph{active} action, as it signifies both involvement and intent to contribute. It should be noted, however, that it only signifies \emph{intent} -- a fork of a repository could be made and no other actions taken.

As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\

@@

The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way commercial software is. ``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}. ``Patches'' in the context of GitHub could be equated to \emph{Pull Requests}. A \emph{Pull Requests} is an announcement by a \emph{fork} that there are changes that may be appropriate to add back to the main repository. The owners of the main repository may then audit these requests, and \emph{pull} some or all of them back into the main repository.

This hypothesis is, ultimately, an examination of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\

@@

\subsection{Research Questions}
To expand the analysis scope slightly, three qualitative research questions are posed. These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub. Where possible, interviews\cite{begel2013social} were conducted to explore high level concepts and motivations.\\

\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A small core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\

@@

\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} computer industry consortium focused on technology standards. Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process. This just one example of how GitHub is being used for non-code artifacts. Are organizations exploring this type of work flow?

\section{Methods}
\subsection{Collection and Storage}
Data collection was done using the GitHub REST API.\footnote{http://developer.github.com/v3/} A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.

\begin{algorithm}[H]
\captionof{algorithm}{Data Collection Algorithm}\label{algo:data_collection_algo}
\begin{algorithmic}[1]
\While{$i < 1000$} 
\State $i \gets i + 1$
\State $word \gets \textsc{random}(word\_list)$
\State $repos \gets \textsc{github\_search}(word)$
\State $j \gets \textsc{random}(repos)$
\State $repo \gets repos_j$
\If { $\textsc{isOrgRepo}(repo)$  }
\State $\textsc{store}(repo)$
\EndIf 
\EndWhile
\end{algorithmic}
\end{algorithm}

Repositories were gathered at random, but only repositories owned by an \textit{Organization} GitHub User type were considered. The \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis. Results were stored in a MySQL relational database management system. The database schema to store the data is described in Figure \ref{fig:er_diagram}.

@@

All source code and materials used in data collection and analysis are hosted at GitHub (naturally!), and may be accessed by the following URL: \url{https://github.com/kevinpeterson/github-process-research}.

\subsection{Analysis}
As stated, the main focus of data collection was to gather metrics pertaining to GitHub \textit{Repositories}. In this context, a \textit{Repository} is defined as a Set of \textit{Attributes}. An \textit{Attribute} is a metric of interest. See Section \ref{sec:attributes} for further information on individual \textit{Attributes}. For analysis purposes, these metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:

@@

All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a set of sets of \textit{Attributes}. Let the set of these sets of \textit{Attributes} be called $S$, or the \textit{Repository Sample}. We can label the \textit{Attribute} sets by arbitrary integers 0 through $|S|$, or:
\[ \{A_i\}_{i \in \{0,1,2 ... |S|\} }\]

In this way, we can reference a set of \textit{Attributes} (or, an individual \textit{Repository}), by $A_i$.

@@

\noindent \textbf{\textit{Commits:}}
The number of individual source code commits to a repository. It is important to note that a \textit{commit} is a single transaction that changes the code contents of a repository. Actual commit ``size'' (lines of code modified, for example) is not taken into account.

@@

\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for several reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests, comments, or general support questions. \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric\cite{phamcreating}.

@@

Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related. The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip. More investigation would be necessary to further examine this finding.\\

\subsection{Research Questions}
\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
The data in Figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown. The intent of this measurement is to show how the number of repository commits are distributed among it's committers. The slices represent the distribution of committers that that have committed the outer percentage amount of commits to a project, or:

\begin{equation}
P = \left\{ \frac{ C_u } { |\mathscr{C}| } \times 100 \Big| u \in U \right\}
\label{eq:commit_percentage}
\end{equation}

Where $U$ is the Set of \textit{Committers} to a given \textit{Repository}, $\mathscr{C}$ is the set of all \textit{Commits} to a \textit{Repository}, and $C_u$ denotes sets of \textit{Commits} by an individual \textit{Committers} to a \textit{Repository}, or:
 \[ \{ C_u \}_{u \in U} \]

Each \textit{Committers}, then, will produce a subset of all \textit{Commits} to a a \textit{Repository}, or $C_u \subset \mathscr{C}$.

\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Distribution of individual committer commit percentage of a repository}
\label{fig:committers_percentage_pie_chart}
\end{figure}

To find the percentages Set $P$ of Equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository. Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range). The distribution of this range assignments is plotted on Figure \ref{fig:committers_percentage_pie_chart}.

Most notably, the data shows that a large percentage of developers committed small amounts ($0-5\%$ of the total repository commit). This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways. Instances of developer monopolization of commits to a repository (for example, where a developer contributes $95-100\%$ of the total commits), are shown to be small.

@@

When looking strictly at the number of committers to a repository, however, the notion of a small set of core committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (Figure \ref{fig:committers_histogram}).\\

\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook. In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.

Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
\begin{quote}
Only in a society where technical supremacy is highly appreciated can developers acquire good reputations among their peers by displaying their skills through free distribution, and often wider acceptance, of their systems.The good reputation attracts attention, trust, and cooperation from others and lays the foundation for advancing the original developers agenda and the establishment and development of OSS communities. 
\end{quote}

With this process change can also come cause for concern. The move to Git from a centralized version control system (such as SVN or CVS) is not without its challenges. Differences in philosophy and nomenclature may place added strain on developers\cite{bird2009promises}. Also, it is not to be assumed that process change is always welcome\cite{de2009software}, even if it is an improvement. Developers and organizations have existing processes, SCM tools, and workflow patterns -- which all change at a cost.

Despite the potential migration cost, the National Center for Biomedical Ontology,\footnote{https://github.com/ncbo}, the National Cancer Informatics Program\footnote{https://github.com/ncip}, NASA\footnote{https://github.com/nasa}, and the WhiteHouse\footnote{https://github.com/WhiteHouse} are just some organizations that have moved all or portions of development to GitHub. As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will move continue to evolve.\\

\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium.\cite{omgTechProcess}
The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.

As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes. Elicitation, Analysis and Negotiation were natural fits for the issue tracker. This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issues tracker itself allowed for dialogs in the form of comments. Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern. OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
Validation and Management are areas of requirements engineering that GitHub has interesting solutions for. First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log. This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue. As the normal OMG\textsuperscript{\textregistered} process proceeds, requirements can be managed through ``tagging'' the issues. By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own work flow. For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration

As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG\textsuperscript{\textregistered} specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.

\section{Conclusion}
The data suggests an evolution in process, but not necessarily a large shift. A small subset of core developers, a traditional aspect of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}. Figure \ref{fig:committers_percentage_pie_chart} shows that a significant number of participation, albeit small, in many repositories. This may indicate a lowered entry barrier for participation.

The data summary variance statement in Assertion: \ref{eq:variance} is important, as it suggests large amounts of variety in the observed variables. This may suggest that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``consumers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises. More research into categorizing GitHub repositories is needed. This is one reason why only repositories owned by a GitHub \textit{Organization} were analyzed, as it was anticipated that this would yield a better sample of repositories for analysis.

The non-code artifacts in GitHub may fall outside of the domain of software development, but it is a significant aspect of our findings. From government documents to software specifications, GitHub is being used and used successfully. More research is needed into what role GitHub will play in the work flow of these artifacts, and whether it will influence their traditional work flow, as it has for software development.
